---
title: "scRNAseq essential analysis with Seurat with sctransform"
author: "Zhuang Liu"
date: "18/09/2021 Updated: `r Sys.Date()`"
output:
  html_document:
    toc: true #add a table of contents (TOC)
    toc_depth: 3 #headers' level in the toc
    toc_float: true #toc visible even when scrolled
    theme: lumen
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', fig.width=12, fig.height = 8)
```

```{r warning=FALSE, message=FALSE}
library(Seurat)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(sctransform)
library(cowplot)
```

#####################
## Reload the data ##
#####################
```{r fig.align='center', fig.width=12, fig.height=7}
hswound.combined.sct <- readRDS("s2_cleanSeurat_harmony_allSamples_clusters_normGenes.rds")
hswound.combined.sct
p1 <- DimPlot(hswound.combined.sct, reduction = "umap", group.by='seurat_clusters', label=TRUE, label.size = 8) +
  ggtitle('Cluster') + NoLegend() + NoAxes()

p1 + FeaturePlot(
  hswound.combined.sct,
  features = c(top20[top20$cluster == 2, ]$gene[1:12]),
  cols = c("gray","red"), ncol = 3
)
```
```{r}
table(hswound.combined.sct$SCT_snn_res.0.8)
markers <- readxl::read_xlsx("s2_MarkerGenes_SingleR_Annotation_manually.xlsx", sheet = 1)
table(markers$cluster)
top20 <- markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_log2FC)
top1 <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_log2FC)
```

## 5.5 Replace the cluster names with annotated cell types
```{r eval=TRUE, fig.align='center', fig.width=12, fig.height=6}
top1$SingleR_main #Check the cell types
####----Annotate the cell types----####
current_cluster_ids <- top1$cluster %>% c() # List of current cluster IDs
new_cluster_ids <- top1$SingleR_main %>% c() # List of new cluster IDs
# Changing IDs to cell type
Idents(hswound.combined.sct) <- plyr::mapvalues(x=Idents(hswound.combined.sct), from=current_cluster_ids, to=new_cluster_ids)
####----Rename the cell types manually----####
#hswound.combined.sct <- RenameIdents(object = hswound.combined.sct,
#                         "0" = "Fibroblasts",
#                         "1" = "Keratinocytes",
#                         "2" = "Keratinocytes",
#                         ...)

####----save general cell type names to metadata and relevel based on order in figures----####
levels(hswound.combined.sct) <- c("Keratinocytes", 
                                  "Fibroblasts", 
                                  "Epithelial cells", 
                                  "Endothelial cells",
                                  "MSCs", #Mesenchymal stem cells 
                                  "CMPs", #common myeloid progenitor cells: can differentiate into granulocyte-macrophage progenitors (GMPs)
                                  "Neurons",
                                  "Dendritic cells",
                                  "NK cells",
                                  "T cells",
                                  "B cells")
hswound.combined.sct$CellTypes_main <- hswound.combined.sct@active.ident

#Draw the umap of annotated cell types
p2 <- DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = TRUE,
        pt.size = 0.5,
        repel = T,
        label.size = 6) + NoAxes()
p1 + p2 + plot_layout(ncol = 2)

####----This step can also reset the active.ident cell names into default number labels----####
hswound.combined.sct <- SetIdent(hswound.combined.sct, value = hswound.combined.sct@meta.data$SCT_snn_res.0.8)
```

## 5.6 AddModuleScore with previously identified cell types
```{r eval=TRUE}
source("../Functions/overlap_phyper2.R")
sfile = "./s2_References for Cell types annotation.xlsx"
sheets = readxl::excel_sheets(sfile)

####----read all sheets and list all the cell types----####
cells.list = list()
for (sheet in sheets) {
    cells.list[[sheet]] = readxl::read_xlsx(sfile, sheet = sheet)
}
#reformat the marker genes from Science paper
cells.list$Science <- pivot_longer(cells.list$Science, everything(), names_to = "cluster", values_to = "gene") %>% arrange(cluster)
  
lapply(cells.list, function(x) unique(x$cluster))

####----select marker genes of our clusters and make a mg.list----####
markers_f <- markers %>% filter(p_val_adj < 0.05) %>% group_by(cluster) %>% top_n(n = 100, wt = avg_log2FC)
mg.list <- split(markers_f$gene, markers_f$cluster)
```

## 5.7.1 celltypes from Science paper
```{r fig.height=6}
cell2gene_1 <- split(cells.list$Science$gene, cells.list$Science$cluster)
overlaps <- overlap_phyper2(cell2gene_1, mg.list)
```
```{r fig.height=60}
ft.names_1 <- paste("Score", names(cell2gene_1), sep = "_")
ft.names_1 <- sub("-", ".", ft.names_1)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_1, name = "CTScores_1", assay = "RNA")
scores_1 <- hswound.combined.sct@meta.data[, grep("CTScores_1", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_1))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_1, col.name = ft.names_1)

VlnPlot(hswound.combined.sct, features = ft.names_1, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```

## 5.7.2 Main celltypes from JACI paper
```{r fig.height=4}
cell2gene_2 <- split(cells.list$All$gene, cells.list$All$cluster)
overlaps <- overlap_phyper2(cell2gene_2, mg.list)
```
```{r fig.height=12}
ft.names_2 <- paste("Score", names(cell2gene_2), sep = "_")
ft.names_2 <- sub("-", ".", ft.names_2)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_2, name = "CTScores_2", assay = "RNA")
scores_2 <- hswound.combined.sct@meta.data[, grep("CTScores_2", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_2))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_2, col.name = ft.names_2)

VlnPlot(hswound.combined.sct, features = ft.names_2, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```

## 5.7.3 Fine celltypes of Keratinocytes from JACI paper
```{r fig.height=4}
####----Keratinocytes----####
cell2gene_3 <- split(cells.list$Keratinocytes$gene, cells.list$Keratinocytes$cluster)
overlaps <- overlap_phyper2(cell2gene_3, mg.list)
```
```{r fig.height=8}
ft.names_3 <- paste("KC_Score", names(cell2gene_3), sep = "_")
ft.names_3 <- sub("-", "_", ft.names_3)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_3, name = "CTScores_3", assay = "RNA")
scores_3 <- hswound.combined.sct@meta.data[, grep("CTScores_3", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_3))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_3, col.name = ft.names_3)

VlnPlot(hswound.combined.sct, features = ft.names_3, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```

## 5.7.4 Fine celltypes of Fibroblasts from JACI paper
```{r fig.height=2}
####----Fibroblasts----####
cell2gene_4 <- split(cells.list$Fibroblasts$gene, cells.list$Fibroblasts$cluster)
overlaps <- overlap_phyper2(cell2gene_4, mg.list)
```
```{r fig.height=5}
ft.names_4 <- paste("FB_Score", names(cell2gene_4), sep = "_")
ft.names_4 <- sub("\\+", "_", ft.names_4)
ft.names_4 <- sub("\\+", "", ft.names_4)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_4, name = "CTScores_4", assay = "RNA")
scores_4 <- hswound.combined.sct@meta.data[, grep("CTScores_4", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_4))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_4, col.name = ft.names_4)

VlnPlot(hswound.combined.sct, features = ft.names_4, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```

## 5.7.5 Fine celltypes of Pericytes, vSMC from JACI paper
```{r fig.height=2}
####----Pericytes, vSMC----####
cell2gene_5 <- split(cells.list$`Pericytes, vSMC`$gene, cells.list$`Pericytes, vSMC`$cluster)
overlaps <- overlap_phyper2(cell2gene_5, mg.list)
```
```{r fig.height=3}
ft.names_5 <- paste("PV_Score", names(cell2gene_5), sep = "_")
ft.names_5 <- sub("-", "_", ft.names_5)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_5, name = "CTScores_5", assay = "RNA")
scores_5 <- hswound.combined.sct@meta.data[, grep("CTScores_5", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_5))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_5, col.name = ft.names_5)

VlnPlot(hswound.combined.sct, features = ft.names_5, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```

## 5.7.6 Fine celltypes of MEL, NC, SC from JACI paper
```{r fig.height=2}
####----MEL, NC, SC----####
cell2gene_6 <- split(cells.list$`MEL, NC, SC`$gene, cells.list$`MEL, NC, SC`$cluster)
overlaps <- overlap_phyper2(cell2gene_6, mg.list)
```
```{r fig.height=3}
ft.names_6 <- paste("MNS_Score", names(cell2gene_6), sep = "_")
ft.names_6 <- sub(" ", "_", ft.names_6)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_6, name = "CTScores_6", assay = "RNA")
scores_6 <- hswound.combined.sct@meta.data[, grep("CTScores_6", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_6))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_6, col.name = ft.names_6)

VlnPlot(hswound.combined.sct, features = ft.names_6, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```

## 5.7.7 Fine celltypes of Sweat gland cells from JACI paper
```{r fig.height=2}
####----Sweat gland cells----####
cell2gene_7 <- split(cells.list$`Sweat gland cells`$gene, cells.list$`Sweat gland cells`$cluster)
overlaps <- overlap_phyper2(cell2gene_7, mg.list)
```
```{r fig.height=3}
ft.names_7 <- paste("SGC_Score", names(cell2gene_7), sep = "_")
ft.names_7 <- sub(" ", "_", ft.names_7)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_7, name = "CTScores_7", assay = "RNA")
scores_7 <- hswound.combined.sct@meta.data[, grep("CTScores_7", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_7))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_7, col.name = ft.names_7)

VlnPlot(hswound.combined.sct, features = ft.names_7, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```

## 5.7.8 Fine celltypes of Vascular endothelial cells from JACI paper
```{r fig.height=3}
####----Vascular endothelial cells----####
cell2gene_8 <- split(cells.list$`Vascular endothelial cells`$gene, cells.list$`Vascular endothelial cells`$cluster)
overlaps <- overlap_phyper2(cell2gene_8, mg.list)
```
```{r fig.height=5}
ft.names_8 <- paste("VE_Score", names(cell2gene_8), sep = "_")
ft.names_8 <- sub("\\/", "_", ft.names_8)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_8, name = "CTScores_8", assay = "RNA")
scores_8 <- hswound.combined.sct@meta.data[, grep("CTScores_8", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_8))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_8, col.name = ft.names_8)

VlnPlot(hswound.combined.sct, features = ft.names_8, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```

## 5.7.9 Fine celltypes of Macrophages, Dendritic cells from JACI paper
```{r fig.height=2}
####----Macrophages, Dendritic cells----####
cell2gene_9 <- split(cells.list$`Macrophages, Dendritic cells`$gene, cells.list$`Macrophages, Dendritic cells`$cluster)
overlaps <- overlap_phyper2(cell2gene_9, mg.list)
```
```{r fig.height=3}
ft.names_9 <- paste("MD_Score", names(cell2gene_9), sep = "_")
ft.names_9 <- sub(" ", "_", ft.names_9)
ft.names_9 <- sub("-", "_", ft.names_9)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_9, name = "CTScores_9", assay = "RNA")
scores_9 <- hswound.combined.sct@meta.data[, grep("CTScores_9", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_9))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_9, col.name = ft.names_9)

VlnPlot(hswound.combined.sct, features = ft.names_9, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```

## 5.7.10 Fine celltypes of T-cells from JACI paper
```{r fig.height=4}
####----T-cells----####
cell2gene_10 <- split(cells.list$`T-cells`$gene, cells.list$`T-cells`$cluster)
overlaps <- overlap_phyper2(cell2gene_10, mg.list)
```
```{r fig.height=5}
ft.names_10 <- paste("T_Score", names(cell2gene_10), sep = "_")
ft.names_10 <- sub("\\+", "", ft.names_10)
ft.names_10 <- sub(" ", "_", ft.names_10)
ft.names_10 <- sub("-", "_", ft.names_10)

# add Celltype scores
hswound.combined.sct <- AddModuleScore(hswound.combined.sct, features = cell2gene_10, name = "CTScores_10", assay = "RNA")
scores_10 <- hswound.combined.sct@meta.data[, grep("CTScores_10", colnames(hswound.combined.sct@meta.data))]
# remove the CTScore columns
hswound.combined.sct@meta.data <- hswound.combined.sct@meta.data[, setdiff(colnames(hswound.combined.sct@meta.data), colnames(scores_10))]

hswound.combined.sct <- AddMetaData(hswound.combined.sct, scores_10, col.name = ft.names_10)

VlnPlot(hswound.combined.sct, features = ft.names_10, group.by='SCT_snn_res.0.8', pt.size = 0, ncol = 3)
```
Export the metadata
```{r}
metadata <- hswound.combined.sct@meta.data %>% tibble::rownames_to_column(var = "CellBarcode")
data.table::fwrite(metadata, "s3_Reference_annotation_celltypeScores.txt", sep="\t")
```

## 5.8 Unsupervised hierarchical clustering results of all clusters
```{r fig.width=12, fig.height=6}
rm(list = ls());gc(verbose = FALSE)
hswound.combined.sct <- readRDS("s1_cleanSeurat_harmony_allSamples_clusters.rds")

Idents(hswound.combined.sct) <- "SCT_snn_res.0.8"
hswound.combined.sct <- BuildClusterTree(object = hswound.combined.sct, assay = "SCT", slot = "scale.data")
levels(hswound.combined.sct) #Check the active cluster levels
suppressMessages(library(scales)) 
show_col(hue_pal()(length(levels(hswound.combined.sct)))) #Check the colors corresponding to each cluster
```
```{r fig.width=4, fig.height=5}
cluster.color <- hue_pal()(length(levels(hswound.combined.sct)))
# pull the tree
hswound.clustertree <- Tool(object = hswound.combined.sct, slot = "BuildClusterTree")
# plot the tree
require(ape)
{plot.phylo(x = hswound.clustertree, direction = "rightwards", use.edge.length = TRUE, font = 4, 
                tip.color = cluster.color,
                edge.width = 2, label.offset = 5, no.margin = TRUE)
tiplabels(pch = 19, col = cluster.color, adj = 3, cex = 2)}
```

```{r}
sessionInfo()
```

