---
title: "Cell type annotations and outlier-cells inspection"
author: "Zhuang Liu"
date: "20/10/3021 Updated: `r Sys.Date()`"
output:
  html_document:
    toc: true #add a table of contents (TOC)
    toc_depth: 3 #headers' level in the toc
    toc_float: true #toc visible even when scrolled
    theme: lumen
    highlight: tango
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', fig.width=12, fig.height = 8)
```

```{r warning=FALSE, message=FALSE}
library(Seurat)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(cowplot)
library(knitr)
library(kableExtra)
```

#####################
# Reload the data
#####################
## 0. Original number labels for each cluster 
```{r fig.width=8, fig.height=6}
hswound.combined.sct <- readRDS("s1_cleanSeurat_harmony_allSamples_clusters.rds") # or read *normGene.rds file
table(hswound.combined.sct$SCT_snn_res.0.8) #cell number for each cluster
table(hswound.combined.sct$orig.ident) #cell number for each sample
table(hswound.combined.sct$Condition)
DefaultAssay(hswound.combined.sct) <- "RNA"

allgene <- rownames(hswound.combined.sct) %>% as.data.frame() %>% setNames("ID")
human.an <- readRDS("../Functions/humanAnnotation.rds")
allgene.f <- allgene %>% left_join(., human.an[,c(1,3)], by=c("ID" = "external_gene_name"))

####----run umap again using other parameters----####
##First save the default umap dimensional reductions as umap_def
#hswound.combined.sct <- RunUMAP(object = hswound.combined.sct, assay = "SCT", reduction = "harmony", dims = 1:40, reduction.name = "umap_def", reduction.key = "UMAPDEF_")

##Second run with optimized parameters
#hswound.combined.sct <- RunUMAP(object = hswound.combined.sct, assay = "SCT", reduction = "harmony", dims = 1:40, n.neighbors = 40, min.dist = 0.3, n.epochs = 200, spread = 1, learning.rate = 1)
#saveRDS(hswound.combined.sct, file = "s1_cleanSeurat_harmony_allSamples_clusters.rds") # or write *normGene.rds file
#library(paletteer) 
#cols <- c(paletteer_d("trekcolors::lcars_series"), "#565656FF", "#E2E2E2FF", "#F8A19FFF")
# define the color
cols <- c(colorRampPalette(c("#1f78b4", "#a6cee3"))(3), #Basal clusters
             colorRampPalette(c( "#33a02c", "#b2df8a"))(4), #Spinous clusters
             "#df65b0", "#d4b9da", #Granular clusters
             colorRampPalette(c("#fdbf6f", "#ff7f00"))(4), #Fibroblast clusters
             "#d9d9d9", "#bc80bd", "#b3de69", "#fb8072", "#fb9a99", #PCvSMC -- VE
             "#fccde5", #Mast cell
             colorRampPalette(c("#cab2d6","#6a3d9a","#ffff99", "#b15928"))(9) #Tcell,myeloid cells
             )
names(ct.cols) <- fact_lev

DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = TRUE, 
        group.by = "SCT_snn_res.0.8", cols = cols) + NoAxes() + NoLegend() + ggtitle("Cell Clusters")
```

## Updated (2022-06-16)
```{r}
kera <- readRDS("../s2_Seurat_subKeratinocytes/allKerat_clean_DelBas4Doublets_InclSpi3_CellID.rds")
fib <- readRDS("../s3_Seurat_subFibroblasts/allFBCells_clean_CellID.rds")
mye <- readRDS("../s4_Seurat_subImmu_MyeloidCells/allMyeloidCells_clean_CellID.rds")
lym <- readRDS("../s5_Seurat_subImmu_LymphoidCells/allLymphoidCells_clean_CellID.rds")
endo <- readRDS("../s6_Seurat_subEndothelial/allEndoCells_clean_CellID.rds")
oth <- readRDS("../s7_Seurat_subOthers/allOtherCells_clean_CellID.rds")

allcleans <- c(kera, fib, mye, lym, endo, oth)

hswound.combined.sct <- hswound.combined.sct[, colnames(hswound.combined.sct) %in% allcleans]
hswound.combined.sct@meta.data <- droplevels(hswound.combined.sct@meta.data)
hswound.combined.sct <- RunUMAP(object = hswound.combined.sct, assay = "SCT", reduction = "harmony", dims = 1:40, reduction.name = "umap_def", reduction.key = "UMAPDEF_")

DimPlot(object = hswound.combined.sct, 
        reduction = "umap_def", 
        label = TRUE, 
        group.by = "SCT_snn_res.0.8") + NoAxes() + NoLegend() + ggtitle("Cell Clusters")

devtools::install_github("EmilHvitfeldt/paletteer")
library(paletteer) 
paletteer_d("Polychrome::light")
library(paletteer) 
paletteer_d("pals::alphabet2")
```


## 1. Annotated cell types for each cluster
```{r}
anno <- readxl::read_xlsx("s5_cell_types_for_each_cluster_manually_updated.xlsx", sheet = 1)
anno %>%
  kbl() %>% 
  kable_material_dark() %>% 
  row_spec(0, bold = T, color = "black") %>%
  scroll_box(width = "100%", height = "400px")
```
```{r eval=TRUE, fig.align='center', fig.width=8, fig.height=6}
####----Annotate the cell types----####
current_cluster_ids <- anno$Cluster %>% c() # List of current cluster IDs
new_cluster_ids <- anno$CellType %>% c() # List of new cluster IDs
# Changing IDs to cell type and add the cell type into metadata
tmp <- plyr::mapvalues(x=Idents(hswound.combined.sct), from=current_cluster_ids, to=new_cluster_ids)
identical(names(tmp), names(Idents(hswound.combined.sct)))
Idents(hswound.combined.sct) <- plyr::mapvalues(x=Idents(hswound.combined.sct), from=current_cluster_ids, to=new_cluster_ids)
##relevel the cell types
hswound.combined.sct$CellTypes <- hswound.combined.sct@active.ident
p1 <- DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = F, 
        group.by = "CellTypes", label.size = 4) + NoAxes() + ggtitle("")
p1

fact_lev <- c("Basal-I", "Basal-II", "Basal-III", "Basal-IV",
              "Spinous-I", "Spinous-II", "Spinous-III",
              "Granular-I", "Granular-II",
              "FB-I", "FB-II", "FB-III", "FB-IV",
              "PC-vSMC", 
              "MEL", 
              "Schwann", 
              "LE", "VE", 
              "Mast-cell", 
              "NK-cell", "Th", 
              "Plasma_B-cell", 
              "Mono-Mac",
              "LC", 
              "DC", "Dermal-DC", "DC-LAMP")

ct.cols <- c(colorRampPalette(c("#1f78b4", "#a6cee3"))(4), #Basal clusters
             colorRampPalette(c( "#33a02c", "#b2df8a"))(3), #Spinous clusters
             "#df65b0", "#d4b9da", #Granular clusters
             colorRampPalette(c("#fdbf6f", "#ff7f00"))(4), #Fibroblast clusters
             "#d9d9d9", "#bc80bd", "#b3de69", "#fb8072", "#fb9a99", #PCvSMC -- VE
             "#fccde5", #Mast cell
             colorRampPalette(c("#cab2d6","#6a3d9a","#ffff99", "#b15928"))(9) #Tcell,myeloid cells
             )
names(ct.cols) <- fact_lev

DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = F, 
        group.by = "CellTypes", label.size = 4, cols = ct.cols) + NoAxes() + ggtitle("")

DefaultAssay(hswound.combined.sct) <- "RNA"
FeaturePlot(hswound.combined.sct, features = c("IFR8", "CLEC9A","CD1C", "LAMP3", "AQP5", "ACTA2"))

wound1 <- subset(hswound.combined.sct, subset = Condition == "Wound1")
wound1$CellTypes <- as.character(wound1$CellTypes)
wound1 <- subset(wound1, subset = CellTypes != "Basal-IV")

wound1$CellTypes <- factor(wound1$CellTypes, levels = fact_lev[-4])
#wound1$CellTypes <- factor(wound1$CellTypes, levels = fact_levs)
p2 <- DimPlot(object = wound1, 
        reduction = "umap", 
        label = F, 
        group.by = "CellTypes", label.size = 4) + NoAxes() + ggtitle("Wound1")
p2
pdf("wound1_umap.pdf", useDingbats = F, width = 8, height = 4)
p2
dev.off()

DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = TRUE, 
        group.by = "CellTypes", label.size = 3) + NoAxes() + ggtitle("Cell Types")
#metadata.output <- hswound.combined.sct@meta.data %>% rownames_to_column(var = "cellID")
#data.table::fwrite(metadata.output, file = "00-Previous_metadata.txt", sep = "\t")

an_maria <- data.table::fread("cell_identity_MariaKasper_Lab.txt", header = FALSE)
colnames(an_maria) <- c("CellID", "Annotated")

mt <- hswound.combined.sct@meta.data %>% rownames_to_column(var = "CellID") %>% 
  left_join(., an_maria, by = c("CellID" = "CellID")) %>% column_to_rownames(var = "CellID")

hswound.combined.sct <- AddMetaData(hswound.combined.sct, metadata = mt)
p1 + DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = TRUE, 
        group.by = "Annotated", label.size = 4) + NoAxes() + NoLegend() + ggtitle("Annotated")
```

## 2. Unsupervised hierarchical tree
```{r fig.width=5, fig.height=6, fig.align='center'}
hswound.combined.sct <- BuildClusterTree(object = hswound.combined.sct, assay = "SCT", slot = "scale.data")
suppressMessages(library(scales))
cluster.color <- hue_pal()(length(levels(hswound.combined.sct)))
# pull the tree
hswound.clustertree <- Tool(object = hswound.combined.sct, slot = "BuildClusterTree")
# plot the tree
suppressMessages(require(ape))
{plot.phylo(x = hswound.clustertree, direction = "rightwards", use.edge.length = TRUE, font = 4, tip.color = cluster.color, edge.width = 2, label.offset = 5, no.margin = TRUE)
tiplabels(pch = 19, col = cluster.color, adj = 3, cex = 2)}

####----This step can also reset the active.ident cell names into default number labels----####
hswound.combined.sct <- SetIdent(hswound.combined.sct, value = hswound.combined.sct@meta.data$SCT_snn_res.0.8)
```


## 3. Proportion of cell types from each batch in each cluster
### 3.1 Plot for each sample
```{r fig.height=8, fig.width=10, warning=FALSE, message=FALSE}
####----Please pay attentions to the original numbers of cells per sample----####
clusters <- unique(hswound.combined.sct@meta.data[["CellTypes"]])
fact_levs <- c("Basal-I", "Basal-II", "Basal-III", "Basal-IV",
                      "Spinous-I", "Spinous-II", "Spinous-III",
                      "Granular-I", "Granular-II",
                      "FB-I", "FB-II", "FB-III", "FB-IV",
                      "PC-vSMC", "MEL", "Schwann", 
                      "LE", "VE", 
                      "Mast-cell", 
                      "NK-cell", "Th", 
                      "Plasma_B-cell", 
                      "Mono-Mac",
                      "LC", 
                      "DC", "Dermal-DC", "DC-LAMP")
levels(clusters) <- rev(fact_levs)

# step 1. 12 individuals divided into 27 cell types and their numbers
step1 <- table(hswound.combined.sct$orig.ident, hswound.combined.sct$CellTypes) %>% as.data.frame()
step1$sepa <- ifelse(step1$Var2 %in% c("Basal-I", "Basal-II", "Basal-III", "Basal-IV",
                                                "Spinous-I", "Spinous-II", "Spinous-III",
                                                "Granular-I", "Granular-II", "LC", "MEL"), "Epidemis", "Dermis")

# step 2. 12 individuals divided into dermis and epidermis and their total cell numbers
step2 <- step1 %>%
  group_by(sepa, Var1) %>%
  summarise(a_sum=sum(Freq)) %>% ungroup() %>% 
  left_join(., step1[, -2:-3]) %>% distinct()
data.table::fwrite(step2, file = "all_epidermis_dermis_perIndi_220613.txt", sep = "\t")

#data.table::fwrite(sepa, file = "allOrig_cellproportion_epidermis_dermis.txt", sep = "\t")

# step 3. calculate the proportion of each cell type for each individual
step3 <- step1 %>% left_join(., step2, by=c("sepa"="sepa", "Var1"="Var1")) %>% distinct() %>% 
  mutate(Prop=Freq/a_sum)

# step 4. calculate the total normalized proportions of each cell type 
step4 <- step3 %>% group_by(Var2) %>% summarise(totalP=sum(Prop))

step5 <- step3 %>% left_join(., step4) %>% mutate(Freq_n = Prop/totalP)

df.group <- step5 %>% mutate(Sample = gsub("PWH..", "", Var1)) %>% group_by(Var2, Sample) %>% 
  summarise(Freq = sum(Freq_n)) %>% ungroup()

df.group$Cluster <- df.group$Var2
df.group$Sample <- gsub("D0", "Skin", df.group$Sample)
df.group$Sample <- gsub("D1", "Wound1", df.group$Sample)
df.group$Sample <- gsub("D7", "Wound7", df.group$Sample)
df.group$Sample <- gsub("D30", "Wound30", df.group$Sample)
df.group$Sample <- factor(df.group$Sample, levels = c("Skin", "Wound1", "Wound7", "Wound30"))

df.group$Condition <- df.group$Sample

df.group$Cluster <- factor(df.group$Cluster, levels = fact_levs)
df.group <- df.group %>% left_join(., anno[, c(2,5)], by=c("Cluster"="CellType"))

df.group.f <- df.group %>% filter(MainCellTypes == "Keratinocytes" & Cluster != "Basal-IV" )

df.group.f <- df.group %>% filter(MainCellTypes == "MastCells" | MainCellTypes == "PlasmaBcells" | MainCellTypes == "PericytesSmoothMuscleCells" | MainCellTypes == "Melanocytes" | MainCellTypes == "SchwannCells")

df.group.f <- df.group

df.group.f$Cluster <- factor(df.group.f$Cluster, levels = fact_levs)

ggplot(df.group.f, aes(x = Cluster, y = Freq, fill = Condition)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_brewer(palette = "Dark2") + # Paired or Set3 maximum variable = 12
  xlab('') +
  scale_y_continuous(breaks = seq(0, 1, 0.1),
                     expand = c(0, 0.01),
                     name = 'Percentage') +
  geom_hline(yintercept=0.25, linetype="dashed", color = "black") +
  geom_hline(yintercept=0.5, linetype="dashed", color = "black") +
  geom_hline(yintercept=0.75, linetype="dashed", color = "black") +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    #panel.border = element_rect(size = 0.35),
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.y = element_text(size = 16, color = "black"),
    legend.position = "right",
    legend.text = element_text(size = 10, color = "black")
  )


clusters <- clusters[order(clusters)]
df <- data.frame()
for(i in 1:length(clusters)){
  SmCell_sum <- table(hswound.combined.sct$orig.ident) %>% as.data.frame()
  #using original cell number of each sample 
  #SmCell_sum <- structure(c(4702, 5035, 4815, 3692, 3597, 7280, 7338, 6605, 5038, 4973, 5973, 6414),
  #                        names=c("PWH26D0", "PWH26D1", "PWH26D7", "PWH26D30", "PWH27D0", "PWH27D1", "PWH27D7", "PWH27D30", "PWH28D0", "PWH28D1", "PWH28D7", "PWH28D30"))
  tmp.df1 <- hswound.combined.sct@meta.data %>% subset(CellTypes == clusters[i]) %>% select(orig.ident) %>% table()
  if(length(tmp.df1) == 12){
    #First normalize the total cell counts per sample
    cur_df <- as.data.frame(tmp.df1 / SmCell_sum)
    colnames(cur_df) <- c("Sample", "Freq")
    #Calculate the normalized proportion 
    cur_df$Freq <- cur_df$Freq * 1/(sum(cur_df$Freq))
    cur_df$Cluster <- clusters[i]
    df <- rbind(df, cur_df)
  } else {
    #print(i);print(tmp.df1)
    #only include the matched samples
    match.sample <- SmCell_sum[names(SmCell_sum) %in% names(tmp.df1)]
    #print(match.sample)
    #First normalize the total cell counts per sample
    cur_df <- as.data.frame(tmp.df1 / match.sample)
    colnames(cur_df) <- c("Sample", "Freq")
    #Calculate the normalized proportion 
    cur_df$Freq <- cur_df$Freq * 1/(sum(cur_df$Freq))
    cur_df$Cluster <- clusters[i]
    df <- rbind(df, cur_df)
  }
}

ggplot(df, aes(x = Cluster, y = Freq, fill = Sample)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_brewer(palette = "Paired") + # Paired or Set3 maximum variable = 12
  xlab('Cell Types') +
  scale_y_continuous(breaks = seq(0, 1, 0.1),
                     expand = c(0, 0),
                     name = 'Percentage') +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    #panel.border = element_rect(size = 0.35),
    axis.text.x = element_text(size = 10, color = "black"),
    axis.text.y = element_text(size = 10, color = "black")
  ) +
  coord_flip()
```

### 3.2 Plot for each condition
```{r fig.height=8, fig.width=10, warning=FALSE, message=FALSE}
df.group <- df %>% mutate(Sample = gsub("PWH..", "", Sample)) %>% group_by(Cluster, Sample) %>% 
  summarise(Freq = sum(Freq)) %>% ungroup()
df.group$Sample <- gsub("D0", "Skin", df.group$Sample)
df.group$Sample <- gsub("D1", "Wound1", df.group$Sample)
df.group$Sample <- gsub("D7", "Wound7", df.group$Sample)
df.group$Sample <- gsub("D30", "Wound30", df.group$Sample)
df.group$Sample <- factor(df.group$Sample, levels = rev(c("Skin", "Wound1", "Wound7", "Wound30")))

df.group$Condition <- df.group$Sample

df.group <- df.group %>% left_join(., anno[,c(2,5)], by=c("Cluster" = "CellType"))
df.group$Cluster <- factor(df.group$Cluster, levels = rev(fact_levs))
ggplot(df.group, aes(x = Cluster, y = Freq, fill = Condition)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_brewer(palette = "Dark2") + # Paired or Set3 maximum variable = 12
  xlab('Cell Types') +
  scale_y_continuous(breaks = seq(0, 1, 0.1),
                     expand = c(0, 0.01),
                     name = 'Percentage') +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    #panel.border = element_rect(size = 0.35),
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 11, color = "black"),
    axis.text.y = element_text(size = 11, color = "black"),
    legend.position = "bottom"
  ) +
  coord_flip() 
```

### 3.3 Plot for each condition (Flip)
```{r fig.height=8, fig.width=10, warning=FALSE, message=FALSE}
####----Please pay attentions to the original numbers of cells per sample----####
smNames_cal <- unique(hswound.combined.sct@meta.data[["Condition"]])
levels(smNames_cal) <- c("Skin", "Wound1", "Wound7", "Wound30")
smNames_cal <- smNames_cal[order(smNames_cal)]

df_sm <- data.frame()
for(i in 1:length(smNames_cal)){
  SmCell_sum <- table(hswound.combined.sct$CellTypes)
  tmp.df1 <- hswound.combined.sct@meta.data %>% subset(Condition == smNames_cal[i]) %>% select(CellTypes) %>% table()
  if(length(tmp.df1) == 27){
    #First normalize the total cell counts per sample
    cur_df <- as.data.frame(tmp.df1 / SmCell_sum)
    colnames(cur_df) <- c("CellTypes", "Freq")
    #Calculate the normalized proportion 
    cur_df$Freq <- cur_df$Freq * 1/(sum(cur_df$Freq))
    cur_df$Sample <- smNames_cal[i]
    df_sm <- rbind(df_sm, cur_df)
  } else {
    #print(i);print(tmp.df1)
    #only include the matched samples
    match.sample <- SmCell_sum[names(SmCell_sum) %in% names(tmp.df1)]
    #print(match.sample)
    #First normalize the total cell counts per sample
    cur_df <- as.data.frame(tmp.df1 / match.sample)
    colnames(cur_df) <- c("CellTypes", "Freq")
    #Calculate the normalized proportion 
    cur_df$Freq <- cur_df$Freq * 1/(sum(cur_df$Freq))
    cur_df$Sample <- smNames_cal[i]
    df_sm <- rbind(df_sm, cur_df)
  }
}

df_sm$CellTypes <- factor(df_sm$CellTypes, levels = 
                            c("Basal-I", "Basal-II", "Basal-III", "Basal-IV",
                              "Spinous-I", "Spinous-II", "Spinous-III",
                              "Granular-I", "Granular-II",
                              "FB-I", "FB-II", "FB-III", "FB-IV",
                              "PC-vSMC", 
                              "MEL", 
                              "Schwann", 
                              "LE", "VE", 
                              "Mast-cell", 
                              "NK-cell", "Th", 
                              "Plasma_B-cell", 
                              "Mono-Mac",
                              "LC", 
                              "DC", "Dermal-DC", "DC-LAMP"))

df_sm <- df_sm %>%
  group_by(Sample, CellTypes) %>%
  mutate(percentage = Freq * 100)
df_sm$TimePoint <- rep(c(0, 1, 7, 30), each=27)
  #mutate(label_y = cumsum(Freq) - 0.1 * Freq) %>%  #Need change the value to adjust the text location
  #mutate(FreqLabel = formatC(Freq * 100,format='f',digits=2)) #Format the label text

cl.celltype <- c('#006d2c', '#31a354', '#74c476', '#a1d99b', 
                 '#08519c', '#3182bd', '#6baed6',
                 '#e6550d', '#fd8d3c',
                 '#756bb1', '#9e9ac8', '#bcbddc', '#dadaeb',
                 '#fb9a99',
                 '#e31a1c',
                 '#fdbf6f',
                 '#878787', '#bababa',
                 '#b15928',
                 '#fee08b', '#ffff99',
                 '#4d4d4d', 
                 '#7fbc41',
                 '#fcc5c0',
                 '#c51b7d', '#de77ae', '#f1b6da')
  
##Plot for each group
ggplot(df_sm, aes(x = Sample, y = percentage, fill = CellTypes)) +
  geom_bar(position = "stack", stat = "identity") +
  #geom_area(alpha=1, size=.3, colour="white") + #Or plot the area barplot
  #scale_fill_igv() + #Change the color
  scale_fill_manual(values = cl.celltype) +
  #geom_text(aes(y = label_y, label = FreqLabel), colour = "white") +
  xlab("Time points") +
  ylab("Cell type proportions") +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    panel.border = element_rect(size = 0.5, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black")
  )
```


## 4. Check the doublet scores and draw the plots
```{r fig.align='center', fig.width=12, fig.height=4}
samples.py <- list.files("../../01_Seurat_qc_Doublets/", pattern = "^Scrublet_")
samples.py.path <- paste0("../../01_Seurat_qc_Doublets/", samples.py)
sm_pyfiles <- lapply(samples.py.path, FUN = function(x){
  data.table::fread(x)
})
smnames <- gsub("Scrublet_", "", samples.py)
smnames <- gsub(".txt", "", smnames)
names(sm_pyfiles) <- smnames

for (i in seq_along(sm_pyfiles)) {
 sm_pyfiles[[i]] <-  sm_pyfiles[[i]] %>% 
   mutate(mad_pre = mad(sm_pyfiles[[i]]$doublet_scores),
          med_pre = median(sm_pyfiles[[i]]$doublet_scores)) %>% 
   mutate(mad_Doublet = ifelse(doublet_scores > (med_pre + 3*mad_pre), "Doublet", "Singlet"))
}
sm_doublets <- do.call("rbind", sm_pyfiles) %>% 
  rename('py_predDoublets' = 'NewDoublets') %>% 
  rename('oldpredicted_doublets' = 'predicted_doublets') %>% select(-1,-5)

##Doublets from pyScrublet (Updted 27 Oct)##
samples.newpy <- list.files("../../01_Seurat_qc_Doublets/pyScrublet/", pattern = ".txt")
samples.newpy.path <- paste0("../../01_Seurat_qc_Doublets/pyScrublet/", samples.newpy)
sm_newpyfiles <- lapply(samples.newpy.path, FUN = function(x){
  data.table::fread(x)
})
names(sm_newpyfiles) <- smnames
for (i in seq_along(sm_newpyfiles)) {
 sm_newpyfiles[[i]] <-  sm_newpyfiles[[i]] %>% 
   mutate(predicted_doublets = ifelse(predicted_doublets == "TRUE", "Doublet", "Singlet")) %>% 
   mutate(barcodename = paste0(smnames[i], "_", barcode))
}
sm_NEWdoublets <- do.call("rbind", sm_newpyfiles) %>% 
  rename('py_NewpredDoublets' = 'predicted_doublets')

####----Doublets from either DoubletFinder or Scrublet----####
sm_doublets_f <- data.table::fread("../../01_Seurat_qc_Doublets/Doublets_either_Scrublet_DoubletFinder.txt")
sm_doublets <- sm_doublets_f %>% select(1, 3) %>% left_join(., sm_doublets, by=c("barcodename" = "barcodename")) %>% 
  left_join(., sm_NEWdoublets[, 3:4], by=c("barcodename" = "barcodename"))

metadata <- hswound.combined.sct@meta.data %>% rownames_to_column(var = "barcode") %>% 
  left_join(., sm_doublets, by=c("barcode" = "barcodename")) %>% 
  mutate(mtPergreat0 = ifelse(percent.mt > 0, "mtGreat0", "mtEqual0")) %>% 
  mutate(mtPer1 = ifelse(percent.mt < 1, "mtLessThan1", "mtGreatThan1")) %>% 
  column_to_rownames(var = "barcode")

metadata_db1 <- metadata %>% filter(predicted_doublets == "Doublet") 
#table(metadata_db1$SCT_snn_res.0.8)
metadata_db2 <- metadata %>% filter(py_predDoublets == "Doublet") 
#table(metadata_db2$SCT_snn_res.0.8)
metadata_db3 <- metadata %>% filter(py_NewpredDoublets == "Doublet") 
#table(metadata_db3$SCT_snn_res.0.8)
metadata_db4 <- metadata %>% filter(mad_Doublet == "Doublet")
#table(metadata_db4$SCT_snn_res.0.8)

metadata_mt <- metadata %>% filter(mtPer1 == "mtLessThan1")
#table(metadata_mt$SCT_snn_res.0.8)
metadata_mt0 <- metadata %>% filter(mtPergreat0 == "mtEqual0")
#table(metadata_mt0$SCT_snn_res.0.8)

hswound.combined.sct <- AddMetaData(hswound.combined.sct, metadata = metadata)
```

### 4.1 Dimplots for checking doublets and mt-percent
```{r fig.align='center', fig.width=12, fig.height=4}
(DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = FALSE, 
        group.by = "predicted_doublets") + NoAxes() + ggtitle("Doublets_either_Two_Tools")) +
  (DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = FALSE, 
        group.by = "py_predDoublets") + NoAxes() + ggtitle("Doublets_only_Srublet")) +
  (DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = FALSE, 
        group.by = "py_NewpredDoublets") + NoAxes())
```
```{r fig.align='center', fig.width=12, fig.height=4}
(DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = FALSE, 
        group.by = "mad_Doublet") + NoAxes()) +
  (DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = FALSE, 
        group.by = "mtPergreat0") + NoAxes()) +
   (DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = FALSE, 
        group.by = "mtPer1") + NoAxes())
```

### 4.2 FeaturePlots for metrics of quality
```{r fig.align='center', fig.width=12, fig.height=8}
# Determine metrics to plot present in hswound.combined.sct@meta.data
metrics <-  c("nCount_RNA", "nFeature_RNA", "S.Score", "G2M.Score", "percent.mt", "doublet_scores")

p3 <- FeaturePlot(
  hswound.combined.sct,
  features = metrics,
  cols = c("gray","red"), combine = FALSE)
p2_list <- lapply(p3, function(x) {x + NoAxes() + theme(text = element_text(size = 10))})
cowplot::plot_grid(plotlist = p2_list, ncol = 3)
```

### 4.3 Violin plots for those metrics of quality
```{r fig.align='center', fig.width=12, fig.height=11}
VlnPlot(hswound.combined.sct, features = metrics, group.by='CellTypes', pt.size = 0, ncol = 2)
```


## 5. Check some known cell type markers
### 5.1 Well-known cell type markers
```{r eval=TRUE, fig.height=4.5, fig.width=12}
####----Marker genes----####
canonical_markers <- list(
  'Keratinocytes' = c('KRT5', 'KRT14', 'KRT10'), 
  'Fibroblasts' = c('COL1A1', 'COL1A2', 'LUM'), 
  'Melanocytes' =	c('PMEL',	'MLANA', 'TYRP1'), 
  'Endothelial-cells' = c('PECAM1', 'VWF', 'EGFL7'), 
  'Myeloid-cells' = c('LYZ', 'HLA-DRA', 'HLA-DRB1'), 
  'Mono-mac' = c('CD68', 'CD163', 'CD14'), #In dermis  
  'Macrophages' = c('CD209', 'FCGR2A', 'MRC1'), #CD209, FCGR2A aka CD32 or FCGR1A aka CD64, 
  #MRC1 aka CD206 (markers of macrophages, immature dendritic cells)
  'T-cells' = c('CD7', 'CD3D', 'CD3E'), 
  'T-cells2' = c('CD4', 'CD8A', 'NKG7'), 
  'B-cells' = c('MS4A1', 'CD79A', 'CD19'), 
  'Plasma-cells' = c('JCHAIN', 'IGHG1', 'MZB1'), 
  'LC-cells' = c('CD207', 'CD1A'), #Primarily located in epidermis (CD1A aka CD1)  
  'Dendritic-cells' = c('CD1C', 'CD1B', 'ITGAX'), #immature DC primarily located in superior dermis (ITGAX aka CD11C)
  'Dendritic-cells2' = c('CD83', 'LAMP3', 'THBD'), #mature DC: LAMP3 aka CD208, THBD aka CD141
  'Dendritic-cells3' = c('CLEC9A'), #CLEC9A aka DNGR1, a Dendritic Cell-Specific Sensor 
  'Schwann' = c('SOX10', 'SOX2'), 
  'PC-vSMC' = c('ACTA2', 'MCAM'), 
  'Epithelial_Mesenchymal' = c('EPCAM', 'PDGFRB')
)
plot_list <- FeaturePlot(
  hswound.combined.sct,
  features=unlist(canonical_markers),
  combine=FALSE, cols = c("gray","red")
)

# apply theme to each feature plot
for(i in 1:length(plot_list)){
  plot_list[[i]] <- plot_list[[i]] + NoLegend() + NoAxes()
}

cowplot::plot_grid(plotlist = plot_list[1:3], ncol = 3) + draw_figure_label(label = "Keratinocytes", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[4:6], ncol = 3) + draw_figure_label(label = "Fibroblasts", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[7:9], ncol = 3) + draw_figure_label(label = "Melanocytes", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[10:12], ncol = 3) + draw_figure_label(label = "Endothelial cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[13:15], ncol = 3) + draw_figure_label(label = "Myeloid Cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[16:18], ncol = 3) + draw_figure_label(label = "Monocytes & Macrophages", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[19:21], ncol = 3) + draw_figure_label(label = "Macrophages", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[22:24], ncol = 3) + draw_figure_label(label = "T/Th cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[25:27], ncol = 3) + draw_figure_label(label = "T/NK cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[28:30], ncol = 3) + draw_figure_label(label = "B cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[31:33], ncol = 3) + draw_figure_label(label = "Plasma cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[34:35], ncol = 3) + draw_figure_label(label = "Langerhans cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[36:38], ncol = 3) + draw_figure_label(label = "Dermal Dendritic cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[39:41], ncol = 3) + draw_figure_label(label = "Mature Dendritic cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[42], ncol = 3) + draw_figure_label(label = "Specific Dendritic cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[43:44], ncol = 3) + draw_figure_label(label = "Schwann cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[45:46], ncol = 3) + draw_figure_label(label = "PC-vSMC cells", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plot_list[47:48], ncol = 3) + draw_figure_label(label = "Epithelial / Mesenchymal cells", size = 8, fontface = "bold", position = "top.right")
```

### 5.2 Cell type markers of basal cells
```{r fig.height=4.5, fig.width=12}
plotcom <- FeaturePlot(
  hswound.combined.sct,
  features=c("KRT14", "KRT10","IVL",
    "ASS1", "COL17A1", "KRT19", "GJB2","PTTG1", "CDC20", "RRM2", "PCLAF", "SPINK5","CALML5"),
  combine=FALSE, cols = c("gray","red")
)
####----markers from https://www.nature.com/articles/s41467-020-18075-7 ----####
#BAS-III cluster (ASS1 and COL17A1) #enrichment between rete ridges
#BAS-III/IV clusters (KRT19 and GJB2) #KRT19 BAS-III show at the bottom/side of the rete ridges, GJB2 at the bottom/side of the rete ridges with some expression in the upper strata 
#BAS-I cluster (PTTG1 and CDC20) #BAS-I and BAS-II basal clusters are enriched for cell cycle marker genes, show a transitional position within the epidermis
#BAS-II cluster (RRM2 and PCLAF)
#GRN cluster (SPINK5 and CALML5)

# apply theme to each feature plot
for(i in 1:length(plotcom)){
  plotcom[[i]] <- plotcom[[i]] + NoLegend() + NoAxes()
}

cowplot::plot_grid(plotlist = plotcom[1:3], ncol = 3) + draw_figure_label(label = "Keratinocytes", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plotcom[4:5], ncol = 3) + draw_figure_label(label = "Basal_rete ridges", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plotcom[6:7], ncol = 3) + draw_figure_label(label = "Basal_rete ridges bottom & upper strata", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plotcom[8:9], ncol = 3) + draw_figure_label(label = "Basal_cell cycle (transitional position)", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plotcom[10:11], ncol = 3) + draw_figure_label(label = "Basal", size = 8, fontface = "bold", position = "top.right")
cowplot::plot_grid(plotlist = plotcom[12:13], ncol = 3) + draw_figure_label(label = "Granular", size = 8, fontface = "bold", position = "top.right")
```

```{r eval=FALSE, fig.height=8, fig.width=12}
hswound.combined.sct <- readRDS("s1_cleanSeurat_harmony_allSamples_clusters.rds") # or read *normGene.rds file
#"FOXM1", "MYBL2", "E2F1", "E2F8", "KLF4", "CEBPA"
plotcom <- FeaturePlot(
  hswound.combined.sct,
  features=c("HOXC13-AS"), #THBD aka CD141, CD1A aka CD1, CD208 aka LAMP3, ITGAX aka CD11C
  combine=F, cols = c("grey","red"), split.by = "Condition"
)

plotcom <- FeaturePlot(
  hswound.combined.sct,
  features=c("HOXC13-AS"), #THBD aka CD141, CD1A aka CD1, CD208 aka LAMP3, ITGAX aka CD11C
  combine=F, cols = c("grey","red")
)

# apply theme to each feature plot
for(i in 1:length(plotcom)){
  plotcom[[i]] <- plotcom[[i]] + NoLegend() + NoAxes()
}

plotcom[[3]] <- DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = TRUE, 
        group.by = "CellTypes", label.size = 3) + NoAxes() + NoLegend() + ggtitle("Cell Types")

cowplot::plot_grid(plotlist = plotcom, ncol = 2)
```

```{r eval=FALSE}
library(rBCS)
hswound.combined.sct <- readRDS("s1_cleanSeurat_harmony_allSamples_clusters.rds") # or read *normGene.rds file
doublet_scores <- data.table::fread("/Users/zhuliu/Desktop/scRNA_STseq/proj_10X_woundhealing/03_results/01-Seurat-PreAnalysis/02_Seurat_BatchCorrection/s1_Seurat_allSample_harmony/py_doublet_OriScores.txt")

doublet_scores_f <- colnames(hswound.combined.sct) %>% as.data.frame() %>% rename("ID" = ".") %>% left_join(., doublet_scores, by=c("ID" = "barcodename"))
hswound.combined.sct$doublet_scores <- doublet_scores_f$doublet_scores

DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = TRUE, 
        group.by = "CellTypes") + NoAxes() + NoLegend() + ggtitle("Cell Clusters")

ExportSeurat(hswound.combined.sct,
             "s2_cleanSeurat_harmony_humanWound_clusters.bcs",
             unique.limit = 100,
             clustering.name = "seurat_clusters",
             compression.level = 5,
             author = "zhuang.liu@ki.se",
             raw.rna = "RNA",
             norm.rna = "RNA",
             overwrite = TRUE)
```

```{r eval=TRUE}

canonical_markers <- c("KRT5", "KRT15", "CENPF", "GJB2", 
                       "KRT10", "KRTDAP", "MT1X", "KRT17",
                       "KRT16", "IVL",
                       "COL1A1", "LUM", "POSTN", "CCL19", "PTGDS", "INHBA",
                       "ACTA2", "MYH11",
                       "TYRP1", "PMEL",
                       "SOX10", "SOX2",
                       "VWF", "CCL21", "SELE",
                       "TPSAB1",
                       "NKG7", "CD3D",
                       "JCHAIN", "MS4A1",
                       "CD14", "CD163", "CD207", "CLEC9A", "CD1C", "LAMP3")

hswound.combined.sct$CellTypes <- factor(hswound.combined.sct$CellTypes, levels = rev(fact_levs))

DotPlot(hswound.combined.sct, features = canonical_markers, 
        group.by = "CellTypes", cols = c("lightgrey", "red"), 
        dot.scale = 4, col.min = 0, dot.min = 0.1
        ) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(x="", y="")


plot_list <- FeaturePlot(
  hswound.combined.sct,
  features=unlist(canonical_markers),
  combine=FALSE, cols = c("gray","red")
)

# apply theme to each feature plot
for(i in 1:length(plot_list)){
  plot_list[[i]] <- plot_list[[i]] + NoLegend() + NoAxes()
}

wrap_plots(plot_list, ncol = 4)
```


```{r eval=TRUE}

canonical_markers <- c("ASPN", "CCL19", "COL13A1", "ANGPTL1",
                       "POSTN", "CXCL3", "APCDD1", "MFAP5",
                       "COL11A1", "APOE", "COL18A1")

plot_list <- FeaturePlot(
  hswound.combined.sct,
  features=unlist(canonical_markers),
  combine=FALSE, cols = c("gray","red")
)

# apply theme to each feature plot
for(i in 1:length(plot_list)){
  plot_list[[i]] <- plot_list[[i]] + NoLegend() + NoAxes()
}

wrap_plots(plot_list, ncol = 4)
```

```{r}
tmp <- FeaturePlot(
  hswound.combined.sct,
  features=c("PCOLCE", "GLIS3"),
  combine=F, cols = c("gray","red"), split.by = "Condition")

for(i in 1:length(tmp)){
  tmp[[i]] <- tmp[[i]] + NoAxes()
}

wrap_plots(tmp, ncol = 2)

DefaultAssay(hswound.combined.sct) <- "RNA"
p1 + (FeaturePlot(
  hswound.combined.sct,
  features=c("LAMB1"),
  combine=T, cols = c("gray","red")
) + scale_color_gradientn(colors = c("grey", "red", "black")))

(FeaturePlot(
  hswound.combined.sct,
  features=c("PLOD2"),
  combine=T, split.by = "Condition", cols = c("gray","red")
) + scale_color_gradientn(colors = c("grey", "red", "black")))
VlnPlot(hswound.combined.sct, features = c("FSIP1")) + NoLegend()
```


# SessionInfo
```{r}
sessionInfo()
```
