---
title: "DE analysis across different conditions"
author: "Zhuang Liu"
date: "07/11/3021 Updated: `r Sys.Date()`"
output:
  html_document:
    toc: true #add a table of contents (TOC)
    toc_depth: 3 #headers' level in the toc
    toc_float: true #toc visible even when scrolled
    theme: lumen
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', fig.width=12, fig.height = 6, warning=FALSE, message=FALSE)
```

```{r warning=FALSE, message=FALSE}
library(Seurat)
library(SeuratObject)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(cowplot)
library(ggrepel)
library(future)
options(future.globals.maxSize = 20000 * 1024^2)
plan("multisession", workers = 4)
```

###################################################
# 1. Preparations of DE analysis across conditions
###################################################
## 1.1 Load the data
```{r warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
hswound.combined.sct <- readRDS("s1_cleanSeurat_harmony_allSamples_clusters.rds")

anno <- readxl::read_xlsx("s5_cell_types_for_each_cluster_manually_updated.xlsx", sheet = 1)

source("../Functions/GOanalysis.R")
anno_human <- readRDS("../Functions/humanAnnotation.rds")

####----DE functions----####
DE.genes <- function(group1 = NULL, group2 = NULL, groupby = NULL, subsetIdent = NULL, minPct = 0.25, fc = 0){
  tmp <- FindMarkers(hswound.combined.sct, 
                     ident.1 = group1, 
                     ident.2 = group2, 
                     group.by = groupby, 
                     subset.ident = subsetIdent,
                     assay = "RNA",
                     test.use = "MAST",
                     latent.vars = c("orig.ident"),
                     min.pct = minPct, 
                     logfc.threshold = fc)
  tmp <- tmp %>% tibble::rownames_to_column(var = "gene") %>% 
    arrange(desc(avg_log2FC)) %>% 
    left_join(., anno_human, by=c("gene" = "external_gene_name")) %>% 
    dplyr::select(1, 3, 2, 6, everything()) %>% 
    #filter(p_val_adj < 0.05) %>% #or filter out the genes with p-value < 0.05
    mutate(Type = factor(ifelse(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3, ifelse(avg_log2FC >= 0.3 ,'Up','Down'),'Not'))) #add a column with regulation direction
  return(tmp)
}

DE.volcano <- function(data = NULL, avgfc = 1, padj = 0.05){
  data.fil <- data %>% filter(abs(avg_log2FC) >= avgfc & p_val_adj < padj)
  ggplot(data=data, aes(x=avg_log2FC, y=-log10(p_val_adj), color=Type, fill=Type)) +
    geom_point(size=1.5)+
    scale_color_manual(values=c("blue","#B3B3B3", "red"))+
    coord_fixed()+
    #geom_vline(xintercept=c(-1,1), lty=4, col="grey", lwd=0.6)+ 
    #geom_hline(yintercept = -log10(0.05), lty=4, col="grey", lwd=0.6)+
    #scale_x_continuous(limits = c(-5,5),breaks = c(-4,-3,-2,-1,0,1,2,3,4)) +
    #scale_y_continuous(limits = c(0,10),breaks = c(0,2,4,6,8,10)) +
    theme_classic() +
    theme(
      panel.grid.major = element_line(colour = "grey90",size = 0.5,linetype = 2),
      panel.grid.minor = element_line(colour = "grey90",size = 0.5,linetype = 2),
      axis.text.x = element_text(size = 12), 
      axis.text.y = element_text(size = 12),
      axis.title.x = element_text(size = 14),
      axis.title.y = element_text(size = 14),
      legend.position = "right",#top,bottom,left,right
      legend.spacing = unit(0,"cm"),
      aspect.ratio = 1) +
    geom_text_repel(data = data.fil, mapping = aes(x=avg_log2FC, y=-log10(p_val_adj), label = gene), show.legend = FALSE,
                    max.overlaps = 50)
}
  
####----Annotate the cell types----####
current_cluster_ids <- anno$Cluster %>% c() # List of current cluster IDs
new_cluster_ids <- anno$CellType %>% c() # List of new cluster IDs
# Changing IDs to cell type and add the cell type into metadata
Idents(hswound.combined.sct) <- plyr::mapvalues(x=Idents(hswound.combined.sct), from=current_cluster_ids, to=new_cluster_ids)
##relevel the cell types
hswound.combined.sct$CellTypes <- hswound.combined.sct@active.ident
table(Idents(hswound.combined.sct))
DimPlot(object = hswound.combined.sct, 
        reduction = "umap", 
        label = TRUE, 
        group.by = "CellTypes", label.size = 3) + NoAxes() + NoLegend() + ggtitle("Cell Types")

```

############################################
# 2. DE analysis of Wound1 vs. Skin
############################################
## 2.1 Keratinocytes
### 2.1.1 Basal-I
```{r fig.width=8, fig.height=8}
celltypename = "Basal-I"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.1.2 Basal-II
```{r fig.width=8, fig.height=8}
celltypename = "Basal-II"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.1.3 Basal-III
```{r fig.width=8, fig.height=8}
celltypename = "Basal-III"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.1.4 Spinous-I
```{r fig.width=8, fig.height=8}
celltypename = "Spinous-I"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.1.5 Spinous-II
```{r fig.width=8, fig.height=8}
celltypename = "Spinous-II"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.1.6 Spinous-III
```{r fig.width=8, fig.height=8}
celltypename = "Spinous-III"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.1.7 Granular-I
```{r fig.width=8, fig.height=8}
celltypename = "Granular-I"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.1.8 Granular-II
```{r fig.width=8, fig.height=8}
celltypename = "Granular-II"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

## 2.2 Fibroblasts
### 2.2.1 FB-I
```{r fig.width=8, fig.height=8}
celltypename = "FB-I"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.2.2 FB-II
```{r fig.width=8, fig.height=8}
celltypename = "FB-II"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.2.3 FB-III
```{r fig.width=8, fig.height=8}
celltypename = "FB-III"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.2.4 FB-IV
```{r fig.width=8, fig.height=8}
celltypename = "FB-IV"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

## 2.3 Melanocytes
### 2.3 MEL
```{r fig.width=8, fig.height=8}
celltypename = "MEL"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

## 2.4 Schwann cells
### 2.4 Schwann
```{r fig.width=8, fig.height=8}
celltypename = "Schwann"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

## 2.5 Pericytes & Vascular smooth muscle cells
### 2.5 PC-vSMC
```{r fig.width=8, fig.height=8}
celltypename = "PC-vSMC"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

## 2.6 Myeloid/Immnue cells
### 2.6.1 Langerhans cells (LC)
```{r fig.width=8, fig.height=8}
celltypename = "LC"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.6.2 Dendritic cells (DC)
```{r fig.width=8, fig.height=8}
celltypename = "DC"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.6.3 Monocytes & Macrophages (Mono-Mac)
```{r fig.width=8, fig.height=8}
celltypename = "Mono-Mac"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.6.4 Mature Dendritic cells (DC-LAMP)
```{r fig.width=8, fig.height=8}
celltypename = "DC-LAMP"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.6.5 Dermal Dendritic cells (Dermal-DC)
```{r fig.width=8, fig.height=8}
celltypename = "Dermal-DC"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.6.6 Mast cells (Mast-cell)
```{r fig.width=8, fig.height=8}
celltypename = "Mast-cell"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.6.7 T help cells (Th)
```{r fig.width=8, fig.height=8}
celltypename = "Th"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.6.8 Natural killer cells (NK-cell)
```{r fig.width=8, fig.height=8}
celltypename = "NK-cell"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.6.9 Plasma & B cells (Plasma_B-cell)
```{r fig.width=8, fig.height=8}
celltypename = "Plasma_B-cell"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

## 2.7 Endothelial cells
### 2.7.1 Vascular endothelial cells (VE)
```{r fig.width=8, fig.height=8}
celltypename = "VE"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

### 2.7.2 Lymphatic endothelial cells (LE)
```{r fig.width=8, fig.height=8}
celltypename = "LE"
DE.genes.sorted <- DE.genes(group1 = "Wound1", group2 = "Skin", groupby = "Condition", subsetIdent = c(celltypename))
DE.volcano(data = DE.genes.sorted)
```

```{r}
####----Filtering and printing all the DE genes----####
DE.genes.sorted %<>% filter(p_val_adj < 0.05 & abs(avg_log2FC) >= 0.3) %>% mutate(Group = "Wound1_Skin")
data.table::fwrite(DE.genes.sorted, file = paste0("s7_Wound1_Skin_", celltypename, "_DEgenes.txt"), sep="\t")

#GO_BP(data = DE.genes.sorted) #Default: using all genes together
GO_BP(data = DE.genes.sorted, regulation = "Up")
GO_BP(data = DE.genes.sorted, regulation = "Down")

#KEGG(data = DE.genes.sorted) #Default: using all genes together
KEGG(data = DE.genes.sorted, regulation = "Up")
KEGG(data = DE.genes.sorted, regulation = "Down")
```

```{r eval=FALSE, fig.width=8, fig.height=8}
library(enrichR)
dbs <- listEnrichrDbs()
dbs <- c("GO_Biological_Process_2021")
enriched <- enrichr(as.character(km_genes[[1]]), dbs)
enriched_bp_up <- enriched$GO_Biological_Process_2021

enriched_d <- enrichr(as.character(unique(defiles.com$`Wound1_Skin_Basal-III` %>% dplyr::filter(Type == "Down") %>% pull(gene))), dbs)
enriched_bp_d <- enriched_d$GO_Biological_Process_2021

ego_BP_up <- enrichGO(gene = defiles.com$`Wound7_Wound1_Basal-I` %>% dylyr::filter(Type == "Up") %>% pull(entrezgene_id),
                          OrgDb         = org.Hs.eg.db,
                          ont           = "BP",
                          keyType = "ENTREZID",
                          pAdjustMethod = "BH",
                          pvalueCutoff  = 0.05,
                          qvalueCutoff  = 0.2,
                          minGSSize = 10,
                          maxGSSize = 500)
dotplot(ego_BP_up, showCategory=10, label_format = 80) + ggtitle("Top 10 GO-BPs") 
```

# SessionInfo
```{r}
sessionInfo()
```
