---
title: "STseq spatial feature plotting"
author: "Zhuang Liu"
date: "8/11/2021 Updated: `r Sys.Date()`"
output: 
  html_document:
    toc: true #add a table of contents (TOC)
    toc_depth: 3 #headers' level in the toc
    toc_float: true #toc visible even when scrolled
    theme: lumen
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', fig.width=12, fig.height = 7, warning=FALSE, message=FALSE)
```

```{r warning=FALSE, message=FALSE}
library(Seurat)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(sctransform)
library(cowplot)
library(harmony)
require(gridExtra)
library(STutility)
```

#####################################################
# Step 1: Spatial Dim Feature Plotting
#####################################################
## 1.1 SpatialDimPlot using Seurat
```{r fig.height=15}
rm(list = ls());invisible(gc())
cl.colors = c('#a6cee3', '#1f78b4', '#b2df8a', '#33a02c', '#fb9a99', '#e31a1c', '#fdbf6f', '#ff7f00', '#cab2d6', '#6a3d9a', '#ffff99', '#b15928',
              '#8dd3c7', '#ffffb3', '#bebada', '#fb8072', '#80b1d3', '#fdb462', '#b3de69', '#fccde5', '#d9d9d9', '#bc80bd', '#ccebc5')
names(cl.colors) = as.character(0:22)

hswound.STcom <- readRDS("s2_Seurat_STseq_integrated_normGenes.rds")
```

#######################################################
# Step 2: TFs Spatial Feature Plotting
#######################################################
## 2.1 Marker genes from curated literatures
```{r eval=TRUE}
canonical_markers <- list(
  'TFs' = c("MMP3", "KRT6A", "KRT17", "KRT6B", "POSTN", "COL1A1")
)

genes <- unlist(canonical_markers)
img.names <- unique(hswound.STcom@meta.data$Sample_ID) %>% as.character()
```


## 2.1 Plotting Donor1_Skin
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor1_Skin", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.1 Plotting Donor1_Wound1
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor1_Wound1", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.1 Plotting Donor1_Wound7
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor1_Wound7", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.1 Plotting Donor1_Wound30
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor1_Wound30", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```


## 2.2 Plotting Donor2_Skin
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor2_Skin", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.2 Plotting Donor2_Wound1
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor2_Wound1", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.2 Plotting Donor2_Wound7
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor2_Wound7", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.2 Plotting Donor2_Wound30
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor2_Wound30", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```


## 2.3 Plotting Donor3_Skin
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor3_Skin", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.3 Plotting Donor3_Wound1
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor3_Wound1", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 4.3 Plotting Donor3_Wound7
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor3_Wound7", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.3 Plotting Donor3_Wound30
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor3_Wound30", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```


## 2.4 Plotting Donor4_Skin
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor4_Skin", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.4 Plotting Donor4_Wound1
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor4_Wound1", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.4 Plotting Donor4_Wound7
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor4_Wound7", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

## 2.4 Plotting Donor4_Wound30
```{r eval=TRUE}
plot_list <- list()
for (i in seq_along(genes)) {
  plot_list[[i]] <- SpatialFeaturePlot(hswound.STcom, features = genes[i],images = "Donor4_Wound30", stroke = 0.4, combine = T, pt.size.factor = 1.7, alpha = c(1, 1), image.alpha = 0.6, crop = TRUE)
}

patchwork::wrap_plots(plot_list[1:3], ncol = 3) + plot_annotation(title = "TFs")
patchwork::wrap_plots(plot_list[4:6], ncol = 3) + plot_annotation(title = "TFs")
```

# Sessioninfo
```{r}
sessionInfo()
```
