---
title: "Slingshot trajectory analysis"
author: "Zhuang Liu"
date: "04/01/2022 Updated: `r Sys.Date()`"
output:
  html_document:
    toc: true #add a table of contents (TOC)
    toc_depth: 3 #headers' level in the toc
    toc_float: true #toc visible even when scrolled
    theme: lumen
    highlight: tango
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', fig.width=12, fig.height = 8, warning=FALSE, message=FALSE)
```

```{r warning=FALSE, message=FALSE}
library(Seurat)
library(tidyverse)
library(magrittr)
library(patchwork)
library(slingshot)
library(tradeSeq)
library(SingleCellExperiment)
library(condiments)
```

###########################################
# Step 1: From Seurat object to slingshot
###########################################
## 1.1 Convert the Seurat object to singlecellexperiment object
```{r}
rm(list = ls());gc()
hswound.krt.com <- readRDS("./allNew_subcluster_keratins_220203.rds") #skin, wound1, wound7, wound30, wounds

DimPlot(object = hswound.krt.com, 
        reduction = "umap", 
        label = TRUE, 
        group.by = "CellTypes") + NoAxes() + ggtitle("Cell Clusters")
DimPlot(object = hswound.krt.com, 
        reduction = "umap", 
        label = TRUE, 
        group.by = "Condition") + NoAxes() + ggtitle("Cell Clusters")

## convert back to singleCellExperiment
sce <- as.SingleCellExperiment(hswound.krt.com, assay = "RNA")
rm(hswound.krt.com);gc()
```

The question is: should we fit a separate trajectory for each condition? We might expect the trajectory itself to be changed by the conditions if the condition effect is systematically large. Otherwise, the condition may impact the expression profile of some genes but the overall trajectory will be preserved.

## 1.2 Balance the overall trajectory
```{r}
df <- bind_cols(
  as.data.frame(reducedDims(sce)$UMAP),
  as.data.frame(colData(sce))
  ) %>% 
  sample_frac(1) %>% rownames_to_column(var = "X1")

sce <- condiments::imbalance_score(Object = sce, 
                                  dimred = "UMAP", 
                                  conditions = "Condition", 
                                  k = 40, smooth = 10)
# k: The number of neighbors to consider when computing the score. Default to 10.
#smooth: The smoothing parameter. Default to k. Lower values mean that we smooth more.

identical(df$X1, colnames(sce[, df$X1]))
df$scores <- sce[, df$X1]$scores$scores
df$scaled_scores <- sce[, df$X1]$scores$scaled_scores

ggplot(df, aes(x = UMAP_1, y = UMAP_2, col = scores)) +
  geom_point() +
  scale_color_viridis_c(option = "C")

ggplot(df, aes(x = UMAP_1, y = UMAP_2, col = scaled_scores)) +
  geom_point() +
  scale_color_viridis_c(option = "C")

ggplot(df, aes(x = UMAP_1, y = UMAP_2, col = CellTypes)) +
  geom_point()
```

## 1.3 Trajectory Inference and Differential Topology (Slingshot)
```{r}
sce <- slingshot(sce, 
                 reducedDim = 'PCA',
                 clusterLabels = sce$CellTypes,
                 start.clus = 'Bas_I', 
                 #approx_points = 150,
                 extend = "n"#,
                 #dist.method = 'mnn'#,
                 #omega = TRUE
                 )#reweight = FALSE, reassign = FALSE

SlingshotDataSet(sce)

set.seed(20220213)
topologyTest(SlingshotDataSet(sce), 
             sce$Condition, 
             rep = 100,
             #methods = "KS_mean", #condition > 2, using Classifier
             threshs = .01)

#method     thresh statistic p.value
#Classifier   0.01      0.99       1
```

```{r}
df$slingPseudotime_1 <- sce[, df$X1]$slingPseudotime_1
df$slingPseudotime_2 <- sce[, df$X1]$slingPseudotime_2
curve <- slingCurves(sce)[[2]] #only see the lineage 1
ggplot(df, aes(x = UMAP_1, y = UMAP_2, col = slingPseudotime_2)) +
  geom_point(size = .7) +
  scale_color_viridis_c() +
  labs(col = "Pseudotime") +
  geom_path(data = curve$s[curve$ord, ] %>% as.data.frame(),
            col = "black", size = 1.5)

```

## 1.5 Differential expression
```{r}
# Differential Progression
progressionTest(sds, conditions = sce$Condition, lineages = TRUE)
# Plot
p5 <- ggplot(df, aes(x = pst)) +
  geom_density(alpha = .4, aes(fill = Condition), col = "transparent") +
  geom_density(aes(col = Condition), fill = "transparent", size = 1.5) +
  guides(col = "none") +
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") +
  labs(x = "Pseudotime", fill = "Type") +
  facet_wrap(~ Curve, scales = "free_x")
p5

# Differential fate selection
fateSelectionTest(sds, conditions = sce$Condition, pairwise = TRUE)
#Plot
weights <- condiments:::.sling_reassign(sds)
df <- df %>%
  full_join(weights %>% 
              as.data.frame() %>%
              mutate(cells = rownames(.)) %>%
              dplyr::rename("Lineage1" = V1, "Lineage2" = V2) %>%
              pivot_longer(starts_with("Lineage"), names_to = "Curve", values_to = "weights")
      )


set.seed(20220213)
filter <- apply(counts(sce), 1, function(g) {
    sum(g >= 5) >= 10
})
sce <- sce[filter, ]

# Select number of knots
set.seed(20220213)
library(BiocParallel)
BPPARAM <- BiocParallel::bpparam()
BPPARAM$workers <- 6
icMat <- evaluateK(counts = as.matrix(assays(sce)$counts),
                   pseudotime = slingPseudotime(sds, na = FALSE),
                   cellWeights = weights,
                   conditions = factor(colData(sce)$Condition),
                   nGenes = 500, #300
                   k = 3:10, #3:7
                   parallel = TRUE,
                   BPPARAM = BPPARAM)

# fit GAM
set.seed(20220213)
sce@int_metadata$slingshot <- sds
sce <- fitGAM(counts = sce,
              conditions = factor(colData(sce)$Condition),
              parallel = TRUE,
              BPPARAM = BPPARAM,
              nknots = 5)
```



# SessionInfo
```{r}
sessionInfo()
```
