---
title: "Stereoscope prediction"
author: "Zhuang Liu"
date: "Updated: `r Sys.Date()`"
output: 
  html_document:
    toc: true #add a table of contents (TOC)
    toc_depth: 3 #headers' level in the toc
    toc_float: true #toc visible even when scrolled
    theme: lumen
    highlight: tango
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align='center', fig.width=12, fig.height = 7, warning=FALSE, message=FALSE)
```

## Load packages
```{r packages}
suppressPackageStartupMessages(require(Matrix))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(Seurat))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(SeuratDisk))
suppressPackageStartupMessages(require(hdf5r))
```

## Prepare and write the scRNA-seq data
```{r}
# load the clean rds object from the step 1 (Figure 1)
sc.data <- readRDS("/Users/zhuliu/Desktop/sc st/plotting scripts/Figure 1/allcombined_wounds_newAnnotation.rds")
DimPlot(sc.data, group.by = "newCellTypes", label = T) + NoAxes()
DimPlot(sc.data, group.by = "newMainCellTypes", label = T) + NoAxes()
```

```{r}
# OBS! All factor columns are converted to numbers in scanpy, fix!
cl = sapply(sc.data@meta.data, is.factor)

for (col in names(cl)[cl]){
    print(col)
    sc.data[[col]] = as.character(sc.data[[col]][,1])
}

sc.data$newMainCellTypes <- gsub(" ", "_",sc.data$newMainCellTypes)
sc.data$newMainCellTypes <- gsub("\\&", "_",sc.data$newMainCellTypes)
sc.data$newMainCellTypes <- gsub("__", "_",sc.data$newMainCellTypes)
sc.data$newMainCellTypes <- gsub("__", "_",sc.data$newMainCellTypes)
table(sc.data$newMainCellTypes)

mt1 <- data.table::fread("/Users/zhuliu/Desktop/sc st/plotting scripts/newsubCluster_annotation/MEL_Schwann.txt") %>% setNames(c("barcode", "upCellTypes"))
mt2 <- data.table::fread("/Users/zhuliu/Desktop/sc st/plotting scripts/newsubCluster_annotation/newFibroblast_metadata.txt")
mt3 <- data.table::fread("/Users/zhuliu/Desktop/sc st/plotting scripts/newsubCluster_annotation/newKeratinocyte_metadata.txt")
mt4 <- data.table::fread("/Users/zhuliu/Desktop/sc st/plotting scripts/newsubCluster_annotation/newMyeloid_metadata.txt")
mt5 <- data.table::fread("/Users/zhuliu/Desktop/sc st/plotting scripts/newsubCluster_annotation/newOther_metadata.txt")
mt6 <- data.table::fread("/Users/zhuliu/Desktop/sc st/plotting scripts/newsubCluster_annotation/newLymphoid_metadata.txt")
mt_all <- rbind(mt1, mt2, mt3, mt4, mt5, mt6) %>% filter(! upCellTypes %in% c("Apoptotic", "Cycling"))
table(mt_all$upCellTypes)

mt_sc <- sc.data@meta.data %>% tibble::rownames_to_column(var = "barcode") %>% select(1,25) %>% setNames(c("barcode", "upCellTypes")) %>% 
  filter(upCellTypes == "Hair_follicle")
mt_all2 <- rbind(mt_all, mt_sc)
mt_all2$upCellTypes <- gsub("-", "_", mt_all2$upCellTypes)
mt_all2$upCellTypes <- gsub("\\(|\\)", "_", mt_all2$upCellTypes)
mt_all2$upCellTypes <- gsub("\\+", "_", mt_all2$upCellTypes)
mt_all2$upCellTypes <- gsub("_$", "", mt_all2$upCellTypes)
mt_all2$upCellTypes <- gsub("_$", "", mt_all2$upCellTypes)
mt_all2$upCellTypes <- gsub("/", "_", mt_all2$upCellTypes)
table(mt_all2$upCellTypes)


sc.data <- DietSeurat(sc.data, assays = "RNA", dimreducs = c("pca", "umap"))
sc.data <- sc.data[, mt_all2$barcode]

mt_all2_f <- sc.data@meta.data %>% tibble::rownames_to_column(var = "barcode") %>% left_join(., mt_all2)
sc.data$upCellTypes <- mt_all2_f$upCellTypes
DimPlot(sc.data, group.by = "upCellTypes", label = T) + NoAxes()


outdir = "/Users/zhuliu/Desktop/scRNA_STseq/proj_spatialTranseq/03_results/deconvolution/cell2location_1118_allsubCT"
# Save as AnnData
SaveH5Seurat(sc.data, filename = file.path(outdir, "scRNAseq_allcleanCells.h5seurat"),
    overwrite = T)
Convert(file.path(outdir, "scRNAseq_allcleanCells.h5seurat"), dest = "h5ad", overwrite = T)

file.remove(file.path(outdir, "scRNAseq_allcleanCells.h5seurat"))
```


# the intermediate steps are same as combined main cell types
```{r}
de.both <- data.table::fread("/Users/zhuliu/Desktop/scRNA_STseq/proj_spatialTranseq/03_results/deconvolution/cell2location_1118_9combCT/scRNAseq_9combCT_allDEs.csv") %>% pull(Gene)

canonical_markers <- c("KRT5", "KRT15", "COL17A1", "MT1G",#Bas-I, Bas-II
                      "CENPF", "MKI67", "MMP3", "MMP1", #Bas-prolif, Bas-mig
                      "KRT10", "MT1X", "DMKN", "LGALS7B", #Spi-I and Spi-II
                      "KRT16", "KRT6A", "FLG", "LORICRIN", #Spi-mig, Gra-I
                      "SOX9", "KRT17", "ADGRL3", #HF
                      "TYRP1", "PMEL", #MEL
                      "COL1A1", "POSTN", "ADAM12", "CCL19", "CXCL12", "PTGDS", "INHBA", #FB-I to IV
                      "SOX10", "SOX2",
                      "ACTA2", "MYH11", "TAGLN", #pc-vsmc
                      "VWF", "PECAM1", "CCL21", "TFF3", "SELE", "ACKR1", #LE, VE
                      "NKG7", "GNLY","CD3D", "TRAC", #NK-T, Th
                      "JCHAIN", "MS4A1", "TPSAB1", "TPSB2", #Bcell, mast-cell
                      "CD163", "CD14", "CLEC9A", "XCR1", "CD1C", "IL1R2", "LAMP3", "FSCN1","CD207", "CD1A")

de.both <- c(de.both, canonical_markers) %>% unique()
```

# load the marker genes from each sub clusters
```{r}
kc <- data.table::fread("/Users/zhuliu/Desktop/scRNA_STseq/proj_10X_woundhealing/03_results/01-Seurat-PreAnalysis/02_Seurat_BatchCorrection/s2_Seurat_subKeratinocytes/00-s2_markergenes_KC_subclusters.txt")
kc_f <- kc %>% mutate(pct.diff = pct.1 - pct.2) %>% filter(avg_log2FC > 0.5) %>% 
  group_by(cluster) %>%
  top_n(90, avg_log2FC) %>% 
  top_n(-60, p_val_adj) %>%
  top_n(40, pct.diff) %>% ungroup() %>% pull(gene) %>% unique()

fb <- data.table::fread("/Users/zhuliu/Desktop/scRNA_STseq/proj_10X_woundhealing/03_results/01-Seurat-PreAnalysis/02_Seurat_BatchCorrection/s3_Seurat_subFibroblasts/s1_clean_fibroblast_mg_res05.txt")
fb_f <- fb %>% mutate(pct.diff = pct.1 - pct.2) %>% filter(avg_log2FC > 0.5) %>% 
  group_by(cluster) %>%
  top_n(90, avg_log2FC) %>% 
  top_n(-60, p_val_adj) %>%
  top_n(40, pct.diff) %>% ungroup() %>% pull(gene) %>% unique()

mye <- data.table::fread("/Users/zhuliu/Desktop/scRNA_STseq/proj_10X_woundhealing/03_results/01-Seurat-PreAnalysis/02_Seurat_BatchCorrection/s4_Seurat_subImmu_MyeloidCells/s1_clean_immu_myeloid_mg_res0.5.txt")
mye_f <- mye %>% mutate(pct.diff = pct.1 - pct.2) %>% filter(avg_log2FC > 0.5) %>% 
  group_by(cluster) %>%
  top_n(90, avg_log2FC) %>% 
  top_n(-60, p_val_adj) %>%
  top_n(40, pct.diff) %>% ungroup() %>% pull(gene) %>% unique()

lym <- data.table::fread("/Users/zhuliu/Desktop/scRNA_STseq/proj_10X_woundhealing/03_results/01-Seurat-PreAnalysis/02_Seurat_BatchCorrection/s5_Seurat_subImmu_LymphoidCells/s1_clean_immu_lymphoid_mg_res0.5.txt")
lym_f <- lym %>% mutate(pct.diff = pct.1 - pct.2) %>% filter(avg_log2FC > 0.5) %>% 
  group_by(cluster) %>%
  top_n(90, avg_log2FC) %>% 
  top_n(-60, p_val_adj) %>%
  top_n(40, pct.diff) %>% ungroup() %>% pull(gene) %>% unique()

oth <- data.table::fread("/Users/zhuliu/Desktop/scRNA_STseq/proj_10X_woundhealing/03_results/01-Seurat-PreAnalysis/02_Seurat_BatchCorrection/s6_7_Seurat_allOthers/s1_clean_othercell_mg_res0.5.txt")
oth_f <- oth %>% mutate(pct.diff = pct.1 - pct.2) %>% filter(avg_log2FC > 0.5) %>% 
  group_by(cluster) %>%
  top_n(90, avg_log2FC) %>% 
  top_n(-60, p_val_adj) %>%
  top_n(40, pct.diff) %>% ungroup() %>% pull(gene) %>% unique()

de.all <- c(de.both, kc_f, fb_f, lym_f, mye_f, oth_f) %>% unique()

write.table(de.all, file = "scRNAseq_allcleanCells_allDEs.csv", row.names = F,
    quote = F, col.names = F)
```


## Write ST data
ST data needs to be loaded into Scanpy from scratch to get positions of images etc correct. So just print out for each sample, the spot ids.

```{r}
st.data = readRDS(file = "/Users/zhuliu/Desktop/sc st/ST plotting scripts/Seurat_STseq_integrate_update.rds")

# Write a csv with barcodes per sample
# Also write the path to the correct input file for each sample.
outdir3="/Users/zhuliu/Desktop/scRNA_STseq/proj_spatialTranseq/03_results/deconvolution/cell2location_1118_9combCT/inputs"
raw.input = "/Users/zhuliu/Desktop/scRNA_STseq/proj_spatialTranseq/03_results/00-spacerange_counts"
subdirs = dir(raw.input, full.names = T)
paths = unlist(sapply(subdirs, dir, full.names = T, pattern = "out"))
snames = unlist(sapply(raw.input, dir, full.names = F))
names(paths) = snames

for(sample in unique(st.data$Sample_name)){
  dir.create(file.path(outdir3,sample), showWarnings = F)
  barcodes = colnames(st.data)[st.data$Sample_name == sample]
  barcodes = unlist(lapply(strsplit(barcodes, "_"), function (x) x[3]))
  write.table(barcodes, file = file.path(outdir3,sample,"barcodes.csv"), row.names = F, quote = F, col.names = F)
  id = unique(st.data$orig.ident[st.data$Sample_name == sample])
  write.table(paths[id], file = file.path(outdir3,sample,"rawdata_path.csv"), row.names = F, quote = F, col.names = F)
}
```

Also write as .mtx together with metadata containing barcode, xcoord, ycoord, 

```{r}
for(sample in unique(st.data$Sample_name)){
  odir = file.path(outdir3,sample)
  C = st.data@assays$Spatial@counts[,st.data$Sample_name == sample]
  writeMM(C, file = file.path(odir, "counts.mtx"))
  write.csv(st.data@images[[sample]]@coordinates, file = file.path(odir,"coordinates.csv"))
}

# also write feature names to a file
write.table(rownames(st.data@assays$Spatial@counts), file = file.path(outdir3, "st_features.csv"), row.names = F, col.names = F)
```

```{r}
# add the metadata
barcode_modi <- readr::read_csv("/Users/zhuliu/Desktop/scRNA_STseq/proj_spatialTranseq/03_results/deconvolution/cell2location_1118_9combCT/inputs/sc_readvisium_barcode_ori.csv")
mt <- st.data@meta.data %>% tibble::rownames_to_column(var ="barcode")

barcode_modi_f <- mt %>% select(1)
readr::write_csv(barcode_modi_f, "/Users/zhuliu/Desktop/scRNA_STseq/proj_spatialTranseq/03_results/deconvolution/cell2location_1118_9combCT/inputs/st_metadata.csv")

vg <- VariableFeatures(st.data, assay = "SCT")
```

### Session info
```{r}
sessionInfo()
```

